\documentclass[11pt]{article}

% ----------- Font Packages ------------
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

% ----------- Math Formatting ----------
\usepackage{amsmath, amssymb, amsthm}
\usepackage{dsfont}
\usepackage[most]{tcolorbox}
\tcbuselibrary{theorems}

% Cross-referencing and hyperlinks
\usepackage{hyperref}
\usepackage{cleveref}

% ----------- Layout and Spacing -------
\usepackage[letterpaper, top=0.75in, bottom=0.75in, left=1in, right=1in, heightrounded]{geometry}
\usepackage{setspace}
\onehalfspacing

% ----------- Misc Enhancements --------
\usepackage{microtype}
\usepackage{enumitem}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhead[L]{Homework One}

\usepackage{bookmark}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    urlcolor=blue,
    citecolor=blue
}

\title{Algebra Homework: Operations with Integers}
\author{Name: \underline{\hspace{3cm}}}
\date{Due: Nov 3}

\begin{document}

\maketitle

% ---------------------- Foundational ----------------------
\section*{Foundational Questions}

\begin{itemize}
    \item What happens when you multiply or divide two negative numbers? How about one negative and one positive? \textit{Hint: multiplying/dividing with negatives flips the number line.}
    \item Why does subtracting a negative number make the result larger? Show me using a real-life example.
    \item Why, when multiplying fractions, do the fractions usually get smaller? Can you tell me scenarios when they would get bigger?
\end{itemize}

% ---------------------- Computation ----------------------
\section*{Integer and Fraction Operations}

Solve the following problems. Show all work and simplify your answers.

\begin{enumerate}[label=\textbf{(\arabic*)}, itemsep=1em]
    \item \( (-3) + 7 \)
    \item \( 5 - (-8) \)
    \item \( (-4) \times 6 \)
    \item \( \dfrac{2}{3} + \dfrac{5}{6} \)
    \item \( \left(-\dfrac{3}{4}\right) - \left(\dfrac{1}{2}\right) \)
    \item \( (-2) \times \left( \dfrac{3}{5} \right) \)
    \item \( \left( \dfrac{4}{7} \right) \div \left( -\dfrac{2}{3} \right) \)
    \item \( (-3) + \left[ 4 - (-2) \times 5 \right] \)
    \item \( \left( \dfrac{3}{4} - \dfrac{1}{8} \right) \times \left( -\dfrac{2}{5} \right) \)
    \item \( \left[ (-6) + \left( \dfrac{3}{2} \right) \times (-4) \right] \div \left( \dfrac{-9}{6} \right) \)
\end{enumerate}

% ---------------------- Word Problems ----------------------
\section*{Word Problems}

Show your work.

\begin{enumerate}[label=\textbf{(\arabic*)}, itemsep=1.5em]
    \item The temperature in Hyde Park yesterday was \(-5^\circ\text{F}\).  
    Today, it increased by \(12^\circ\text{F}\).  
    What is the temperature now?

    \item A scuba diver is \(40\) feet below sea level.  
    She rises \(18\) feet, then descends another \(7\) feet.  
    What is her final position relative to sea level?

    \item Abdullah spent \(\$ \dfrac{3}{4}\) of his money on a book and \(\$ \dfrac{1}{8}\) on a snack.  
    What fraction of his money does he have left?

    \item A football team gained \(15\) yards on its first play, lost \(22\) yards on the second, and gained \(9\) yards on the third.  
    What is the team’s total change in yards?

    \item The temperature dropped \(\dfrac{3}{5}\) degrees every hour for \(4\) hours.  
    What was the total change in temperature?

    \item \textbf{Challenge Problem}

    In a Fortnite match, a player starts with \(-120\) shield points (he’s taken heavy storm damage).  
    He uses a shield potion that restores \(\dfrac{2}{3}\) of his lost shield.  
    Then, he takes an additional \(45\) points of fall damage.  
    Finally, he finds a legendary item that restores \(\dfrac{1}{2}\) of his current shield amount.  

    \begin{enumerate}[label=\textbf{(\alph*)}]
        \item Write an expression using integers, fractions, and parentheses to represent the player’s final shield amount.  
        \item Simplify the expression step by step to find his ending shield points.
    \end{enumerate}
\end{enumerate}


\end{document}
